---
title: "HA2_DA3"
author: "Bokonyi Gabor; Illes Gabor; Radics Gabor"
date: "26/11/2017"
output:
  pdf_document: default
  html_document: default
---


## Problem #2.1
```{r, echo=FALSE}
setwd("/Volumes/Mac_500GB_SSD/IG_Docs/CEU/Lectures/DA_3/HA_test")
library(wbstats)
library(data.table)
library(ggplot2)
library(lspline)
library(rmarkdown)
```

## 1. Data acquisition and cleaning
#### Download cros-county 'GDP per capita' data
```{r}
GDP <- wb(indicator = "NY.GDP.PCAP.PP.KD", startdate = 2011, enddate = 2011)
```
#### Change coloumn names
```{r}
setnames(GDP, 'value', 'GDP_per_capita')
setnames(GDP, 'country', 'country_name')
```
#### Download cros-county 'Life Expectancy' data
```{r}
LEX <- wb(indicator = "SP.DYN.LE00.IN", startdate = 2011, enddate = 2011)
```
#### Change coloumn names
```{r}
setnames(LEX, 'value', 'life_expectancy')
```
#### Merge the two datasets by 'iso2c' code
```{r}
P21 <- data.table(merge(GDP[, c('iso2c', 'country_name', 'GDP_per_capita')], LEX[, c('iso2c', 'life_expectancy')], by = 'iso2c', all.x = TRUE, all.y = TRUE))
```
#### Drop observations with missing values and write csv file
```{r}
P21 <- P21[(GDP_per_capita != 'NA' & life_expectancy != 'NA'),]
write.csv(P21[, c('country_name', 'life_expectancy', 'GDP_per_capita')], 'P21.csv')
```

## 2. Estimate regressions
#### Histogram of 'Life Expectancy'
```{r, echo=FALSE}
ggplot(P21, aes(x = life_expectancy)) + geom_histogram(binwidth = 1, colour = "darkgreen", fill ="darkgreen") + ggtitle("life expectancy (ys)")
```
some explanation here

#### Histogram of 'GDP per capita'
```{r, echo= FALSE}
ggplot(P21, aes(x = GDP_per_capita)) + geom_histogram(binwidth = 1000, colour = "darkgreen", fill ="darkgreen") + ggtitle("GDP per capita")
```
some explanation here

```{r, echo=FALSE}
P21[, `:=`(
  GDP_per_capita_M = GDP_per_capita / 1000, 
  GDP_per_capita_M_sq = (GDP_per_capita / 1000) ^ 2,
  GDP_per_capita_M_cub = (GDP_per_capita / 1000) ^ 3,
  GDP_per_capita_log = log(GDP_per_capita)
),]
```

#### Lowess regression of 'Life Expectancy' on ln 'Gdp per capita'

```{r, echo=FALSE}
ggplot(P21, aes(GDP_per_capita_log, life_expectancy)) + 
  geom_point(size = 1.5, colour="orange", shape = 4) + 
  geom_smooth(colour = "darkgreen") + 
  labs(x="GDP per capita - log", y="life expectancy (ys)") +
  ggtitle("lowess regression")
```
some explanation here

#### Summary statistics of Lowess regression of 'Life Expectancy' on ln 'Gdp per capita'
```{r, echo=FALSE}
#ls <- lowess(life_expectancy ~ GDP_per_capita_log, data = P21)
#ls2 <- lowess(P21$GDP_per_capita_log, P21$life_expectancy)
#summary(ls2)
```


#### Simple Linear Regression of 'Life Expectancy' on 'Gdp per capita'

```{r, echo=FALSE}
slr <- lm(life_expectancy ~ GDP_per_capita_M, data = P21)

ggplot(data = slr, aes(x=GDP_per_capita_M, y=life_expectancy)) +
  geom_point(size=1.5, colour="orange",shape=4) +
  geom_smooth(method="lm", colour="darkgreen", se=FALSE)+
  labs(x="GDP per capita (M$)",y="life expectancy (ys)")+
  ggtitle("simple linear regression")
```
some explanation here


##### Summary statistics of 'Life Expectancy' on 'Gdp per capita'
```{r, echo=FALSE}
summary(slr)
```
some explanation here


#### Level-log linear regression of 'Life Expectancy' on 'Gdp per capita'

```{r, echo=FALSE}
levellog <- lm(life_expectancy ~ GDP_per_capita_log, P21)

ggplot(data = P21, aes(x=GDP_per_capita_log, y=life_expectancy)) +
  geom_point(size=1.5, colour="orange") +
  labs(x="GDP per capita - log",y="life expectancy (ys)")+
  geom_smooth(method="lm", colour="darkgreen", se=F)+
  ggtitle("level-log linear regression")
```
some explanation here

##### Summary statistics of level-log regression of 'Life Expectancy' on 'Gdp per capita'
```{r, echo=FALSE}
summary(levellog)
```
some explanation here



#### Piecewise linear spline of 'Life Expectancy' on 'Gdp per capita' (lspline)

```{r, echo=FALSE}
pls <- lm(life_expectancy ~ lspline(GDP_per_capita_M, c(25)), data=P21)
P21$pls<-predict(pls)

ggplot(data = P21, aes(x=GDP_per_capita_M, y=life_expectancy)) +
  geom_point(size=1.5, colour="orange",shape=4) +
  labs(x="GDP per capita (M$)", y="life expectancy (ys)")+
  geom_line(data = P21,aes(x=GDP_per_capita_M, y=pls),colour="darkgreen")+
  geom_vline(xintercept=25,colour="red")+
  ggtitle("piecewise linear spline")
```
some explanation here

##### Summary statistics of linear spline of 'Life Expectancy' on 'Gdp per capita'
```{r, echo=FALSE}
summary(pls)
```
some explanation here


#### Quadratic regression of 'Life Expectancy' on 'Gdp per capita'

```{r, echo=FALSE}
quad <- lm(life_expectancy ~ GDP_per_capita_M + GDP_per_capita_M_sq, P21)
P21$pred_quad<-predict(quad)

ggplot(data = P21, aes(x=GDP_per_capita_M, y=life_expectancy)) +
  geom_point(size=1.5, colour="orange",shape=4) +
  labs(x="GDP per capita (M$)", y="life expectancy (ys)")+
  geom_line(data=P21, aes(x=GDP_per_capita_M,y=pred_quad), colour="darkgreen")+
  ggtitle("quadratic model")




```
some explanation here

##### Summary statistics of Quadratic regression of 'Life Expectancy' on 'Gdp per capita'
```{r, echo=FALSE}
summary(quad)
```
some explanation here


#### Cubic regression of 'Life Expectancy' on 'Gdp per capita'

```{r, echo=FALSE}
cub <- lm(life_expectancy ~ GDP_per_capita_M + GDP_per_capita_M_sq + GDP_per_capita_M_cub, P21)
P21$pred_cub<-predict(cub)

ggplot(data = P21, aes(x=GDP_per_capita_M, y=life_expectancy)) +
  geom_point(size=1.5, colour="orange",shape=4) +
  labs(x="GDP per capita (M$)", y="life expectancy (ys)")+
  geom_line(data=P21, aes(x=GDP_per_capita_M, y=pred_cub), colour="darkgreen")+
  ggtitle("cubic model")




```
some explanation here

##### Summary statistics of Cubic regression of 'Life Expectancy' on 'Gdp per capita'
```{r, echo=FALSE}
summary(cub)
```
some explanation here


## 3. Estimate a weighted regression 

```{r, include=FALSE}
wbsearch(pattern = "Population, total")
POP <- wb(indicator = "SP.POP.TOTL", startdate = 2011, enddate = 2011)
setnames(POP, 'value', 'population')
PS21w <- data.table(merge(P21, POP[, c('iso2c', 'population')], by = 'iso2c'))
PS21w$population_MM <- PS21w$population / 1000000




```

#### weighted level-log linear regression


```{r, echo=FALSE}
w_levellog <- lm(life_expectancy ~ GDP_per_capita_log, data = PS21w, weights = population_MM)

ggplot(data = PS21w, aes(x=GDP_per_capita_log, y=life_expectancy, size=population_MM)) +
  geom_point(colour="orange") +
  labs(x="GDP per capita - log",y="life expectancy (ys)")+
  geom_smooth(method="lm", colour="darkgreen", se=F)+
  ggtitle("weighted level-log linear regression")




```
some explanation here


```{r, echo=FALSE}
summary(w_levellog)




```
some explanation here







```{r, echo=FALSE}




```


```{r, echo=FALSE}




```









